Supplementary material to: Empirical Bayes estimation of the normal inverse Gaussian prior in simultaneous-equation models

### 2019-08-13 docs/supplement1.Rnw
updated simulations and manuscript

	\section{NIG prior}\label{sec:nigprior}
	For a fixed error variance $\sigma_d^2$, we have a marginal prior variance 
	and kurtosis:
	\begin{align*}
	  \V(\beta_{jd} | \sigma_d^2) & = \sigma_d^2 \theta_d, \\
	  \mathcal{K}(\beta_{jd} | \sigma_d^2) & = \frac{3\theta_d}{\lambda_d} + 3.
	\end{align*}

	\section{NIGIG prior}\label{sec:nigigprior}
	Equivalently, we may set 
  $\eta_{jd}^2=\tau_j^2 \gamma_d^2/(\theta_j^z \theta_d^c)$ and write 
  MD equation (\ref{md-eq:generalmodel}) as:
  \begin{align*}
    \beta_{jd} | \eta_{jd}^2, \sigma_d^2 & \sim \mathcal{N} 
    \(0, \frac{\lambda_{jd}}{\chi_{jd}}\eta_{jd}^2\sigma_d^2\), \\
    \eta_{jd}^2 & \overset{D}{=}\pi^{-1}\sqrt{\chi_{jd}}\exp(\psi_{jd}) 
    \eta_{jd}^{-3/2} \cdot K_{0} \(\sqrt{\psi_{jd}^2 + \chi_{jd} (\eta_{jd}^2 
    + \eta_{jd}^{-2} - 2)} \),
  \end{align*}
  where $\lambda_{jd}=\lambda_j^z\lambda_d^c$, 
  $\chi_{jd}=\lambda_j^z\lambda_d^c/(\theta_j^z \theta_d^c)$, and
  $\psi_{jd}=\lambda_j^z/\theta_j^z + \lambda_d^c/ \theta_d^c$. A third way
  of writing the model is:
  \begin{subequations}\label{eq:generalprior}
    \begin{align}
      \beta_{jd} | \eta_{jd}^2, \sigma_d^2 & \sim \mathcal{N} 
      \(0, \eta_{jd}^2 \sigma_d^2\), \\
      \eta_{jd}^2 | \xi_{d}^2 & \sim \mathcal{IG} 
      (\theta_j^z \xi_d^2, \lambda_j^z \xi_d^2 ), \\
      \xi_d^2 & \sim \mathcal{GIG} (1, \lambda_d^c/(\theta_d^c)^2, \lambda_d^c).
    \end{align}
  \end{subequations}
	From (\ref{eq:generalprior}), for a fixed error variance
	$\sigma_d^2$, we find the marginal prior variance and kurtosis:
	\begin{subequations}\label{eq:varkurtnigig}
	  \begin{align}
	    \V(\beta_{jd} | \sigma_d^2) & = \sigma_d^2 \theta_d^c \theta_j^z  
      \(r_d + \frac{2\theta_d^c}{\lambda_d^c}\), \\
	    \mathcal{K}(\beta_{jd} | \sigma_d^2) & = \(\frac{3\theta_j^z}
	    {\lambda_j^z} + 3\) \frac{4 \theta_d^c (r_d + 2\theta_d^c/\lambda_d^c) +
	    1}{\lambda_d^c(r_d + 2\theta_d^c/\lambda_d^c)^2},
	  \end{align}
	\end{subequations}
	with $r_d = K_0(\lambda_d^c/\theta_d^c)/K_1(\lambda_d^c/\theta_d^c)$.
	If 
  we fix the prior expectation $\E(\gamma_d^2)=\theta_d^c$ and let the prior 
  variance 
  $\V(\gamma_d^2)$ go to zero, we have $\V(\beta_{jd} | \sigma_d^2) \rightarrow
  \sigma_d^2 \theta_j^z \theta_d^c$ and $\mathcal{K}(\beta_{jd} | \sigma_d^2)
  \rightarrow 3(\theta_j^z/\lambda_j^z + 1)$. From (\ref{eq:varkurtnigig}) we see
  that we can keep the variance constant by fixing $\theta_j^z$, $\theta_d^c$,
  and $\lambda_d^c$, while we increasing the kurtosis by decrearing 
  $\lambda_j^z$. This makes the model MD equations (\ref{md-eq:generalmodel}) a
  flexible model.

	\section{Student's \texorpdfstring{$t$}{t} prior}\label{sec:studentstprior}
	For a fixed error variance $\sigma_d^2$, we have a marginal prior variance 
	and kurtosis:
	\begin{align*}
	  \V(\beta_{jd} | \sigma_d^2) & = \frac{\sigma_d^2 \lambda_d}{\eta_d - 2}, \\
	  \mathcal{K}(\beta_{jd} | \sigma_d^2) & = \frac{6}{\eta_d - 4} + 3.
	\end{align*}

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	For clarity we have indexed the variational 
	density functions with their respective parameters, which we omitted in the 
	MD. In the following all expectations are with respect to the 
	variational posterior $Q_d$. The variational posterior for $\bbeta_d$ is 
	found as follows:
	\begin{align*}
	  \log q_{\bbeta_d} (\bbeta_d) & \propto \E [\log \mathcal{L}
	  (\y_d | \bbeta_d, \sigma_d^2 )] + \E [\log \pi (\bbeta_d | \gamma_d^2, 
	  \sigma_d^2)] \\
	  & \propto - \frac{1}{2} \sum_{i=1}^n 
	  \E \[ \frac{(y_{id} - \x_i \tr \bbeta_d)^2}{\sigma_d^2} \] - 
	  \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_d^2 \sigma_d^2} \), \\
	  q_{\bbeta_d} (\bbeta_d) & \overset{D}{=} 
	  \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	  & \text{with } \bSigma_d = \E(\sigma_d^{-2})^{-1} 
	  [\X \tr \X + \E(\gamma_d^{-2}) \cdot \I ]^{-1}, \\
	  & \text{and } \bmu_d = [\X \tr \X + \E(\gamma_d^{-2}) \cdot \I ]^{-1} 
	  \X \tr \y_d.
	\end{align*}
	The variational posterior for $\gamma_d^2$ is given by:
	\begin{align*}
	  \log q_{\gamma_d^{2}}(\gamma_d^{2}) & \propto \E 
	  [\log \pi (\bbeta_d | \gamma_d^{2}, \sigma_d^2)] + \log \pi(\gamma_d^{2}) \\
	  & \propto -\frac{p}{2} \log \gamma_d^{2} - \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_d^{-2} - 
	  \frac{3}{2} \log \gamma_d^{2} - 
	  \frac{\lambda_d (\gamma_d^2 - \theta_d)^2}{2 \theta_d^2} \gamma_d^{-2} \\
	  & \propto \( -\frac{p + 1}{2} - 1\) \log \gamma_d^{2} - 
	  \frac{\lambda_d}{2\theta_d^2} \gamma_d^2 - 
	  \frac{1}{2}\[ \lambda_d + \E(\sigma_d^{-2}) \sum_{j=1}^p \E(\beta_{jd}^2)\] 
	  \gamma_d^{-2} , \\
	  q_{\gamma_d^{2}}(\gamma_d^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-\frac{p+1}{2}, \frac{\lambda_d}{\theta_d^2}, \delta_{d} \), \\
	  & \text{with } \delta_{d} =\E(\sigma_d^{-2})
	  \left\{\E(\bbeta_{d} \tr) \E(\bbeta_{d}) + \trace[\V(\bbeta_{d})]\right\} + 
	  \lambda_d.
	\end{align*}
	Lastly, we have the variational posterior for $\sigma_d^2$:
	\begin{align*}
	  \log q_{\sigma_d^{2}}(\sigma_d^{2}) & \propto \E 
	  [\log \mathcal{L}(\y_d | \bbeta)d, \sigma_d^2 )] + 
	  \E [\log \pi (\bbeta_d | \gamma_d^2, \sigma_d^2)] + 
	  \log \pi (\sigma_d^{2}) \\
	  & \propto -\frac{n}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{i=1}^n 
	  \frac{\E [(y_{id} - \x_i \tr \bbeta_d)^2]}{\sigma_d^{2}} - 
	  \frac{p}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2)}{\gamma_d^{-2}}\) \sigma_d^{-2} - 
	  \frac{3}{2} \log \sigma^2_d \\
	  & = \(-\frac{n + p + 1}{2} - 1 \) \log \sigma_d^{2} - 
	  \frac{1}{2}\left\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d)^2] + 
	  \E(\gamma_d^{-2})\sum_{j=1}^p \E(\beta_{jd}^2) \right\} \sigma_d^{-2}, \\
	  q_{\sigma_d^{2}}(\sigma_d^{2}) & \overset{D}{=} 
	  \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_{d} \), \\
	  & \text{with } \zeta_{d} = \frac{1}{2} \bigg\{\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \E(\bbeta_d) + 
    \trace [ \X \tr \X \V(\bbeta_d)] + 
    \E(\bbeta_d \tr)  \X \tr \X \E(\bbeta_d ) \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \E(\gamma_d^{-2}) \trace [ \V(\bbeta_d)] + 
    \E(\gamma_d^{-2}) \E(\bbeta_d \tr) \E(\bbeta_d ) \bigg\}.
	\end{align*}

	\section{Evidence lower bound}\label{sec:evidencelowerbound}
	The variational Bayes posterior is found through maximisation of an evidence
	lower bound (ELBO). Additionally, we may monitor convergence through this 
	ELBO. To compute the ELBO after updating the EB parameters from 
	$\theta_d^{(l)}, \lambda_d^{(l)}$ to 
	$\theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$, we compute 
	$\text{ELBO}^{(l + 1)} = \sum_{d=1}^D \text{ELBO}_d^{(l + 1)}$, with:
	\begin{align*}
	  \text{ELBO}_d^{(l + 1)} = & \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - 
	  \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + 
	  (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + 
	  \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
	  & + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - 
	  a_d^{(l)}(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - 
	  a_d^{(l)} \trace (\bm{\Sigma}_d^{(l)}) \] \\
	  & - \frac{p + n + 1}{2} \log \zeta_d^{(l)} + 
	  \frac{p + 1}{4} \log \lambda_d^{(l)} - 
	  \frac{p + 1}{2} \log \theta_d^{(l)} - \frac{p+1}{4} \log \delta_d^{(l)} \\
	  & + \log K_{\frac{p+1}{2}}\(\sqrt{\lambda_d^{(l)} 
	  \delta_d^{(l)}}/\theta_d^{(l)}\) + 
	  \frac{\lambda_d^{(l)} e_d^{(l)}}{2(\theta_d^{(l)})^2} \\
	  & + \frac{\lambda_d^{(l+1)}}{\theta_d^{(l+1)}} + 
	  \frac{1}{2}\log \lambda_d^{(l+1)} - \frac{\lambda_d^{(l+1)} 
	  e_d^{(l)}}{2 (\theta_d^{(l + 1)})^2} - 
	  \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2},
	\end{align*}
	where $a^{(l)}$, $b^{(l)}$, and $e^{(l)}$ are as in the MD. To compute the 
	ELBO after a VB update we simply let 
  $\theta_d^{(l)}, \lambda_d^{(l)}=\theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$.

  In practice, there are different options to assess convergence. We may monitor
  the parameters themselves, the ELBO, or a combination of the two. We may also 
  iterate for an \textit{a priori} fixed number of times. We may also choose to 
  use a different convergence criterium for the EB and VB iterations. Currently 
  we fix the number of EB iterations to 20, while we do 2 VB iterations per EB 
  iteration.

  \section{Efficient computation}
  The empirical Bayes updates require the following quantities: $\zeta_d$, $a_d$, $\delta_d$, $\trace (\bm{\Sigma}_d)$, $\trace (\X \tr \X \bm{\Sigma}_d)$, $\bm{\mu}_d \tr \bm{\mu}_d$, and $\bm{\mu}_d \tr \X \tr \X \bm{\mu}_d$. In addition, we need $\log|\bm{\Sigma}|$ and $\mathbf{y}_d \tr \X \bm{\mu}_d$ to monitor the ELBO. The first three quantities are obtained by scalar operations of $\mathcal{O}(n)$. Let $c_d^{(h)}=b_d^{(h)}$ in the conjugate setting and $c_d^{(h)}=b_d^{(h)}/a_d^{(h)}$ in the non-conjugate setting. Then, beforementioned quantities are easily calculated using the SVD $\X = \mathbf{U} \mathbf{D} \mathbf{V} \tr$:
  \begin{align*}
  \trace (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} \[\sum_{i=1}^n (v_i^2 + c_d^{(h)})^{-1} + \max (p - n, 0)\cdot (c_d^{(h)})^{-1}\], \\
  \trace (\X \tr \X \bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} \sum_{i=1}^n v_i^2(v_i^2 + c_d^{(h)})^{-1}, \\
  (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)})^2, \\
  (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^4 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)})^2, \\
  \log|\bm{\Sigma}^{(h+1)}| & \propto -p \log a_d^{(h)} - \sum_{i=1} \log (v_i^2 + c_d^{(h)}) - \max(p - n, 0) \cdot \log c_d^{(h)}, \\
  \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)}).
  \end{align*}
  where $v_i$ are the singular values of $\X$. The SVD and $\mathbf{U} \tr \mathbf{y}_d$ are $\mathcal{O}(pn^2)$ and $\mathcal{O}(n^2)$ operations, respectively, but have to be calculated only once at the start of the algorithm. The rest of the calculations involve just $n$ scalar multiplications and/or additions.

  \section{Ratios of modified Bessel functions}\label{sec:ratiosmodifiedbessels}
  Ratios of modified Bessel functions of the second kind 
  $K_{\alpha - 1}(x)/K_{\alpha}(x)$ are prone to under- and overflow for large
  $\alpha$. In our case, $\alpha$ increases linearly with $p$. Since $p$ may be
  large, this causes numerical issues in the calculation of various quantities.
  We alleviate the numerical issues through the following.

@@

	\section{Non-conjugate model}
	The non-conjugate model assumes $\beta_{jd}$ and $\sigma_d^2$ 
	independent \textit{a priori}:
  $$
  \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_d^2 \sim \mathcal{N} (0, \gamma_d^2).
  $$

	The variational distributions are as follows:
	\begin{align*}
	  q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
	  (\bm{\mu}_d, \bm{\Sigma}_d), \\
	  q(\gamma_d^2) & \overset{D}{=} \mathcal{GIG}
	  \(-\frac{p+1}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
	  q(\sigma_d^2) & \overset{D}{=} \Gamma^{-1} 
	  \(\frac{n + 1}{2}, \zeta_d\),
  \end{align*}
	The variational parameters are iteratively updated by:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} (\X \tr \X + c_d^{(h)} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + c_d^{(h)} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = b_d^{(h)} (c_d^{(h)})^{-1} \bmu_d \tr \bmu_d + \lambda_d, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{b_d^{(h+1)} - c_d^{(h+1)} a_d^{(h+1)}}{1 - a_d^{(h+1)}} \trace ( \bm{\Sigma}_d^{(h+1)}) + \frac{b_d^{(h+1)} - c_d^{(h+1)} a_d^{(h+1)}}{1 - a_d^{(h+1)}} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg].
  \end{align*}
  Here, we set $a_d^{(h)}$, $b_d^{(h)}$, and $c_d^{(h)}$ as in 
  (\ref{eq:auxvar1})
  Note that in the inverse Gamma model, the computation of (\ref{eq:auxvar1})
  simplifies to $b_d^{(h)}=(p+\eta_d)/\delta_d^{(h)}$.

>>>>

	\section{NIG and Student's \texorpdfstring{$t$}{t} Priors}\label{sec:priors}
	\subsection{Normal inverse Gaussian}\label{sec:nigprior}
	Let $\theta = (\mathbf{c} \tr \bm{\alpha})^{-1}$. Then, 
	for a fixed error variance $\sigma^2$, we have marginal prior variance 
	and kurtosis:
	\begin{align*}
	  \V(\beta | \sigma^2) & = \sigma^2 \theta, \\
	  \mathcal{K}(\beta | \sigma^2) & = \frac{3\theta}{\lambda} + 3.
	\end{align*}

	\subsection{Student's \texorpdfstring{$t$}{t} prior}\label{sec:studentstprior}
	For a fixed error variance $\sigma^2$, we have marginal prior variance 
	and kurtosis:
	\begin{align*}
	  \V(\beta | \sigma^2) & = \frac{\sigma^2 \xi}{\eta - 2}, \\
	  \mathcal{K}(\beta | \sigma^2) & = \frac{6}{\eta - 4} + 3.
	\end{align*}

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	For clarity we have indexed the variational 
	density functions with their respective parameters, which we omitted in the 
	MD. In the following all expectations are with respect to the 
	variational posterior $Q_d$. The variational posterior for $\bbeta_d$ is 
	found as follows:
	\begin{align*}
	  \log q_{\bbeta_d} (\bbeta_d) & \propto \E [\log \mathcal{L}
	  (\y_d | \bbeta_d, \sigma_d^2 )] + \E [\log \pi (\bbeta_d | \bgamma_{d}^2, 
	  \sigma_d^2)] \\
	  & \propto - \frac{1}{2} \sum_{i=1}^n 
	  \E \[ \frac{(y_{id} - \x_i \tr \bbeta_d)^2}{\sigma_d^2} \] - 
	  \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_{jd}^2 \sigma_d^2} \), \\
	  q_{\bbeta_d} (\bbeta_d) & \overset{D}{=} 
	  \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	  & \text{with } \bSigma_d = \E(\sigma_d^{-2})^{-1} 
	  \{\X \tr \X + \diag [\E(\gamma_{jd}^{-2})] \}^{-1}, \\
	  & \text{and } \bmu_d = \{\X \tr \X + \diag [\E(\gamma_{jd}^{-2})] \}^{-1} 
	  \X \tr \y_d.
	\end{align*}
	The variational posterior for $\gamma_{jd}^2$ is given by:
	\begin{align*}
	  \log q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \propto \E 
	  [\log \pi (\beta_{jd} | \gamma_{jd}^{2}, \sigma_d^2)] + 
	  \log \pi(\gamma_{jd}^{2}) \\
	  & \propto -\frac{1}{2} \log \gamma_{jd}^{2} - \frac{1}{2} 
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_{jd}^{-2} - 
	  \frac{3}{2} \log \gamma_{jd}^{2} - 
	  \frac{\lambda_d \{\gamma_{jd}^2 - 
	  [(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}\}^2}
	  {2[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-2}} \gamma_{jd}^{-2} \\
	  & \propto \( -1 - 1\) \log \gamma_{jd}^{2} - 
	  \frac{\lambda_d[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{2}}{2} \gamma_{jd}^2 - 
	  \frac{1}{2}\[ \lambda_d + \E(\sigma_d^{-2}) \E(\beta_{jd}^2)\] 
	  \gamma_{jd}^{-2} , \\
	  q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-1, \lambda_d[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{2}, \delta_{jd} \), \\
	  & \text{with } \delta_{jd} =\E(\sigma_d^{-2})
	  \[\E(\beta_{jd})^2 + \V(\beta_{jd})\] + \lambda_d.
	\end{align*}
	Lastly, we have the variational posterior for $\sigma_d^2$:
	\begin{align*}
	  \log q_{\sigma_d^{2}}(\sigma_d^{2}) & \propto \E 
	  [\log \mathcal{L}(\y_d | \bbeta)d, \sigma_d^2 )] + 
	  \E [\log \pi (\bbeta_d | \bgamma_d^2, \sigma_d^2)] + 
	  \log \pi (\sigma_d^{2}) \\
	  & \propto -\frac{n}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{i=1}^n 
	  \frac{\E [(y_{id} - \x_i \tr \bbeta_d)^2]}{\sigma_d^{2}} - 
	  \frac{p}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2)}{\gamma_{jd}^{2}}\) \sigma_d^{-2} - 
	  \frac{3}{2} \log \sigma^2_d \\
	  & = \(-\frac{n + p + 1}{2} - 1 \) \log \sigma_d^{2} - 
	  \frac{1}{2}\left\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d)^2] + 
	  \sum_{j=1}^p \E(\gamma_{jd}^{-2})\E(\beta_{jd}^2) \right\} \sigma_d^{-2}, \\
	  q_{\sigma_d^{2}}(\sigma_d^{2}) & \overset{D}{=} 
	  \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_{d} \), \\
	  & \text{with } \zeta_{d} = \frac{1}{2} \bigg\{\mathbf{y}_d \tr 
	  \mathbf{y}_d - 2 \mathbf{y}_d \tr \X \E(\bbeta_d) + 
    \trace [ \X \tr \X \V(\bbeta_d)] + 
    \E(\bbeta_d \tr)  \X \tr \X \E(\bbeta_d ) \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \trace \{ \diag[\E(\gamma_{jd}^{-2})] \V(\bbeta_d)\} + 
     \E(\bbeta_d \tr) \diag[\E(\gamma_{jd}^{-2})] \E(\bbeta_d ) \bigg\}.
	\end{align*}

% 	\section{Evidence lower bound}\label{sec:evidencelowerbound}
% 	The variational Bayes posterior is found through maximisation of an evidence
% 	lower bound (ELBO). Additionally, we may monitor convergence through this 
% 	ELBO. To compute the ELBO after updating the EB parameters from 
% 	$\theta_d^{(l)}, \lambda_d^{(l)}$ to 
% 	$\theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$, we compute 
% 	$\text{ELBO}^{(l + 1)} = \sum_{d=1}^D \text{ELBO}_d^{(l + 1)}$, with:
% 	\begin{align*}
% 	  \text{ELBO}_d^{(l + 1)} = & \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - 
% 	  \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + 
% 	  (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + 
% 	  \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
% 	  & + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - 
% 	  a_d^{(l)}(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - 
% 	  a_d^{(l)} \trace (\bm{\Sigma}_d^{(l)}) \] \\
% 	  & - \frac{p + n + 1}{2} \log \zeta_d^{(l)} + 
% 	  \frac{p + 1}{4} \log \lambda_d^{(l)} - 
% 	  \frac{p + 1}{2} \log \theta_d^{(l)} - \frac{p+1}{4} \log \delta_d^{(l)} \\
% 	  & + \log K_{\frac{p+1}{2}}\(\sqrt{\lambda_d^{(l)} 
% 	  \delta_d^{(l)}}/\theta_d^{(l)}\) + 
% 	  \frac{\lambda_d^{(l)} e_d^{(l)}}{2(\theta_d^{(l)})^2} \\
% 	  & + \frac{\lambda_d^{(l+1)}}{\theta_d^{(l+1)}} + 
% 	  \frac{1}{2}\log \lambda_d^{(l+1)} - \frac{\lambda_d^{(l+1)} 
% 	  e_d^{(l)}}{2 (\theta_d^{(l + 1)})^2} - 
% 	  \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2},
% 	\end{align*}
% 	where $a^{(l)}$, $b^{(l)}$, and $e^{(l)}$ are as in the MD. To compute the 
% 	ELBO after a VB update we simply let 
%   $\theta_d^{(l)}, \lambda_d^{(l)}=\theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$.
% 	 
%   In practice, there are different options to assess convergence. We may monitor
%   the parameters themselves, the ELBO, or a combination of the two. We may also 
%   iterate for an \textit{a priori} fixed number of times. We may also choose to 
%   use a different convergence criterium for the EB and VB iterations. Currently 
%   we fix the number of EB iterations to 20, while we do 2 VB iterations per EB 
%   iteration.

  \section{Efficient computation}
  The empirical Bayes updates require the quantities
  $e_{jd}$ and $b_{jd}$, which in turn depend on the quantities 
  $\diag(\bSigma_d)$, $\bmu_{d}$, $\y_d \tr \X \bmu_d$, 
  $\trace(\X \tr \X \bSigma_d)$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $\trace[\diag(b_{jd})\bSigma_d]$, and $\bmu_d \tr \diag(b_{jd}) \bmu_d$.
  Let $\mathbf{H}_d=\diag[(b_{jd}^{(h)})^{-1}]$. Then, the first two quantities 
  are more efficiently calculated as:
  \begin{subequations}\label{eq:complexity1}
    \begin{align}
      \diag (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} 
      \H_d^{(h)} - (a_d^{(h)})^{-1} \{ [ \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} ] \circ
      (\H_d^{(h)} \X \tr ) \} \mathbf{1}_{n \times 1}, \\
      \bmu_d^{(h+1)}) & = \H_d^{(h)} \X \tr \y_d - \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr \y_d.
    \end{align}
  \end{subequations}
  The quantities $\y_d \tr \X \bmu_d$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $\trace[\diag(b_{jd})\bSigma_d]$, and $\bmu_d \tr \diag(b_{jd}) \bmu_d$ are
  easily calculated from the first two. For the remaining one we have:
  \begin{align}\label{eq:complexity2}
    \trace(\X \tr \X \bSigma_d^{(h+1)})  = 
    & (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \H_d^{(h)} \X \tr 
    \mathbf{1}_{n \times 1} \nonumber \\  
    & - (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \{ [  \H_d^{(h)} \X \tr 
    ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr ] \circ 
    \X \tr \} \mathbf{1}_{n \times 1}.
  \end{align}
  Both (\ref{eq:complexity1}) and (\ref{eq:complexity2}) are operations of
  complexity $\mathcal{O}(pn^2)$.

  \section{Ratios of modified Bessel functions}\label{sec:ratiosmodifiedbessels}
  Ratios of modified Bessel functions of the second kind 
  $K_{\alpha - 1}(x)/K_{\alpha}(x)$ are prone to under- and overflow for large
  $\alpha$. In our case, $\alpha$ increases linearly with $p$. Since $p$ may be
  large, this may cause numerical issues in the calculation of various 
  quantities. We alleviate the numerical issues through the following.

@@

	\section{Log pseudo marginal likelihood}\label{sec:lpml}
	The log pseudo marginal likelihood is calculated as:
	$\text{LPML}_d = n^{-1}\sum_{i=1}^n \log (\text{CPO}_{id})$ 
  \cite[]{gelfand_model_1992}. Let 
  $\theta_{id}=\x_i \tr \bbeta_d$ and $\y_{-id}$ denote $\y_{d}$ with 
  observation $y_{id}$ excluded. Then the conditional 
  predictive ordinates (CPO) are given by:
  \begin{align*}
    \text{CPO}_{id} & = p(y_{id} | \y_{-id}) = 
    \int_{-\infty}^{\infty} \int_0^{\infty}
    p(y_{id} | \bbeta_d, \sigma_d^2 ) \pi (\bbeta_d | \y_{-id}, \sigma_d^2) 
    \pi (\sigma_d^2 | \y_{-id}) d\sigma_d^2 d\bbeta_d \\
    & = \int_{-\infty}^{\infty} \int_0^{\infty}
    p(y_{id} | \theta_{id}, \sigma_d^2 ) 
    \pi (\theta_{id} | \y_{-id}, \sigma_d^2) 
    \pi (\sigma_d^2 | \y_{-id}) d\sigma_d^2 d\theta_{id},
  \end{align*}
  where
  \begin{align}
    y_{id} | \theta_{id}, \sigma_d^2 & \sim \mathcal{N} (\theta_{id}, 
    \sigma_d^2), \nonumber \\
    \theta_{id} | \y_{-id} & \sim \mathcal{N}(\x_i \tr \bmu_{-id}, 
    \x_i \tr \bSigma_{-id} \x_i), \nonumber \\
    \sigma_d^2 | \y_{-id} & \sim \Gamma^{-1} \( \frac{n + p}{2},
    \zeta_{-id} \), \label{eq:gammavb}
  \end{align}
  where $\bmu_{-id}$, $\bSigma_{-id}$, and $\zeta_{-id}$ are the variational
  parameters estimated from $\y_{-id}$. Plugging in the densities we see that 
  we have:
  \begin{subequations}
    \begin{align}
      \text{CPO}_{id} = & (2 \pi)^{-1} \frac{\Gamma\(\frac{n + p + 1}
      {2}\)}{\Gamma\(\frac{n + p}{2}\)}
      (\zeta_{-id} \x_i \tr \bSigma_{-id} \x_i)^{-1/2} \label{eq:cpo1} \\
      & \cdot \int_{-\infty}^{\infty}
      \( \frac{\theta^2}{2\zeta_{-id}} + 1 \)^{-\frac{n + p + 1}{2}}
      \exp \[ - \frac{(\theta - \x_i \tr \bmu_{-id} + y_i)^2}
      {2 \x_i \tr \bSigma_{-id} \x_i} \] d\theta, \label{eq:cpo2}
    \end{align}
  \end{subequations}
  a one-dimensional (well-behaved) integral that is easily solved with some 
  numerical integration routine.

  In practice we would like to avoid re-estimating the model for every 
  observation $i$. Instead we divide the observations into $K$ parts and 
  estimate $K$ models, so that now $\y_{-id}$ denotes $\y$ with all observations
  in part $k$ removed. Additionally, we add $1 - n_{id}$ to the numerator of the
  Gamma distribution in (\ref{eq:gammavb}), the numerators in the Gamma 
  functions in (\ref{eq:cpo1}), and the numerator of the power in
  (\ref{eq:cpo2}), where $n_{id}$ denotes the number of elements in the removed 
  part.

% 	\section{Non-conjugate model}
% 	The non-conjugate model assumes $\beta_{jd}$ and $\sigma_d^2$ 
% 	independent \textit{a priori}:
%   $$
%   \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | 
%   \gamma_d^2 \sim \mathcal{N} (0, \gamma_d^2).
%   $$
%   
% 	The variational distributions are as follows:
% 	\begin{align*}
% 	  q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
% 	  (\bm{\mu}_d, \bm{\Sigma}_d), \\
% 	  q(\gamma_d^2) & \overset{D}{=} \mathcal{GIG}
% 	  \(-\frac{p+1}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
% 	  q(\sigma_d^2) & \overset{D}{=} \Gamma^{-1} 
